---
title: "Peptidomique_clivages_final"
author: "Marion Hardy"
date: "2023-04-13"
output: 
  html_document:
    toc: true 
    theme: spacelab 
    highlight: monochrome
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, cache = TRUE, echo = FALSE, warning = F, cache.lazy = F)

library(tidyverse)
library(readxl)

```

```{r data loading}

data = read_xlsx("./data/Peptidomics36_peptides_site clivage.xlsm")
prot = read_xlsx("./data/BVDE_929_data.xlsx", sheet = "PeptideGroups")
list = read_xlsx("./data/Peptidomics36_peptides_site clivage.xlsm", sheet = "enriched_pep")

```

## Proteomic data

First, we will filter those proteomics data.

```{r, include=TRUE, echo=TRUE}

prot = 
  prot %>% 
  filter(Contaminant != "TRUE")

# Reduce the number of rows to keep only the ones of interest

prot = prot[,c(3:5,9,10:11,16:23,29:44)]
colnames(prot)[11:14] = c("F1_H2O2","F1_ctrl","F2_H2O2","F2_ctrl")

# Filter out peptides which appear in <4 conditions

cols = sapply(prot[15:30], grepl, pattern = "Not.*{2,}")
prot = prot[rowSums(cols) >= 3, ]

# Add the "oxidated" and "not oxidated" annotations

prot =
prot %>%
  mutate(state = case_when(
    grepl("xidation", Modifications) ~ "Oxidated",
    !grepl("xidation", Modifications) ~ "Not oxidated"))


```


```{r, include=F}

# If we wanted to keep only the enriched in H2O2 and ctrl. 
prot_f =
  prot %>% 
  filter(`Abundance Ratio: (F1, H2O2) / (F1, Non R)` != 1 |
           `Abundance Ratio: (F2, H2O2) / (F2, Non R)` != 1,
         `Abundance Ratio Adj. P-Value: (F2, H2O2) / (F2, Non R)`<= 0.05|
            `Abundance Ratio Adj. P-Value: (F1, H2O2) / (F1, Non R)`<= 0.05)

length(unique(prot_f$`Annotated Sequence`))

table(prot_f$`Abundance Ratio: (F2, H2O2) / (F2, Non R)`>1)

```

# Peptidomic data

Peptidomics experiment: Eluate peptides ( = epitopes) from IgG and analyze through MS.

Sophie has around 140 enriched peptides in her peptidomics analysis. 
It's based on Nathalie's Log2Fc on the peptide abundances (from the peptidomic, not the proteomic data) and is filtered on pval < 0.05. 

```{r}
head(data)[1:9] %>% 
  knitr::kable()
```

data\$\`Amino acid before\` I'm guessing this is the aa before the cleavage
data\$\`Amino acid after\` this is the aa following the cleaved sequence

## Check the aa at Cter and/or Nter

```{r, include=TRUE, echo=TRUE}

list = 
  data %>% 
  filter(Sequence%in%list$Enriched_pep)

pep = 
  list %>% 
    filter(`Amino acid before`%in%c("M","C")|
            `Amino acid after`%in%c("M","C")|
             `First amino acid`%in%c("M","C")|
             `Last amino acid`%in%c("M","C"))

nrow(pep)

pep %>% 
  select(Sequence, `Amino acid before`,`Amino acid after`,`First amino acid`, `Last amino acid`,Proteins, `Oxidation (M) site IDs`) %>% 
  knitr::kable()

```

21/140 enriched peptides contain
- M or C in Nter
- and/or M or C in Cter

Just by curiosity, how many are found to be oxidized?

```{r, echo=TRUE, include=TRUE}

pepox =
  pep %>% 
    filter(!is.na(`Oxidation (M) site IDs`))

pepox %>% 
  select(Sequence, `Amino acid before`,`Amino acid after`,`First amino acid`, `Last amino acid`,Proteins, `Oxidation (M) site IDs`) %>% 
  knitr::kable()

```

Only four of these peptide is found oxidized in the peptidomics data.

## Check if the proteins of those 21 peptides can be found in the proteomics data.


```{r, include=TRUE, echo=TRUE}
library(stringr)

pep =
  pep %>%
  mutate(Uniprot = str_match(pep$Proteins, "(^tr|sp)\\|(.*?)\\|")[,3])

common =  
  prot %>% 
  filter(`Master Protein Accessions`%in%pep$Uniprot)

dim(common)

length(unique(common$`Annotated Sequence`))

length(unique(common$`Master Protein Accessions`))

common %>%
  select(`Annotated Sequence`,Modifications, `Master Protein Accessions`, `Abundance Ratio: (F2, H2O2) / (F2, Non R)`,
         `Abundance Ratio Adj. P-Value: (F2, H2O2) / (F2, Non R)`,state) %>% 
  head(50) %>% 
  knitr::kable()

```

There are 190 (200 total) unique peptides from 11/21 proteins that were significantly enriched in H2O2 or control whose peptides can also be found to have a cystein or a methionine next to the cleave site.

## Which are those proteins?

```{r}

uni_pep = unique(common$`Master Protein Accessions`)

pep %>%
  filter(Uniprot%in%uni_pep) %>% 
  select(Sequence, `Amino acid before`,`Amino acid after`,`First amino acid`, `Last amino acid`,Proteins, `Oxidation (M) site IDs`) %>% 
  knitr::kable()

```


## In which state are they found?

```{r, echo=TRUE, include=TRUE}

table(common$state)

```

## In which condition are they found?

```{r, echo=TRUE, include=TRUE}

table(common$state, common$`Abundance Ratio: (F2, H2O2) / (F2, Non R)`>1)

```

- 83 non oxidated enriched in control condition
- 9 oxidated encriched in control condition
- 54 non oxidated found in H2O2
- 17 oxidated found in H2O2


## To which oxidated protein correspond those oxidated peptides found in the proteomics data

```{r, include=TRUE}

H2O2enriched = 
common %>% 
  filter(`Abundance Ratio: (F2, H2O2) / (F2, Non R)`>1,
         state == "Oxidated") %>% 
  select(`Master Protein Accessions`)

ctrlenriched = 
common %>% 
  filter(`Abundance Ratio: (F2, H2O2) / (F2, Non R)`<1,
         state == "Oxidated") %>% 
  select(`Master Protein Accessions`)

uni_pepH2O2 = unique(H2O2enriched$`Master Protein Accessions`)
uni_pepctrl = unique(ctrlenriched$`Master Protein Accessions`)


```

### Oxidated in H2O2

```{r}
pep %>%
  filter(Uniprot%in%uni_pepH2O2) %>% 
  select(Sequence, `Amino acid before`,`Amino acid after`,`First amino acid`, `Last amino acid`,Proteins, `Oxidation (M) site IDs`) %>% 
  knitr::kable()
```

### Oxidated in control condition

```{r}
pep %>%
  filter(Uniprot%in%uni_pepctrl) %>% 
  select(Sequence, `Amino acid before`,`Amino acid after`,`First amino acid`, `Last amino acid`,Proteins, `Oxidation (M) site IDs`) %>% 
  knitr::kable()
```

## Taking a look at those peptides in the proteomics data:

### Oxidated in H2O2

```{r}

aa =
  pep %>%
    filter(Uniprot%in%uni_pepH2O2)

prot %>% 
  filter(`Master Protein Accessions`%in% aa$Uniprot,
         !is.na(`Abundance Ratio Adj. P-Value: (F2, H2O2) / (F2, Non R)`)) %>% 
  group_by(`Master Protein Accessions`, state) %>% 
  summarise(`Master Protein Accessions`, state, meanH2O2_ctrl = mean(`Abundance Ratio: (F2, H2O2) / (F2, Non R)`)) %>% 
  distinct() %>% 
  knitr::kable()

```

### Oxidated in ctrl

```{r}
bb =
  pep %>%
    filter(Uniprot%in%uni_pepctrl)

prot %>% 
  filter(`Master Protein Accessions`%in% bb$Uniprot,
         !is.na(`Abundance Ratio Adj. P-Value: (F2, H2O2) / (F2, Non R)`)) %>% 
  group_by(`Master Protein Accessions`, state) %>% 
  summarise(`Master Protein Accessions`, state, meanH2O2_ctrl = mean(`Abundance Ratio: (F2, H2O2) / (F2, Non R)`)) %>% 
  distinct() %>% 
  knitr::kable()

```

